---
title: "STAT 156 Final Project ReAnalysis"
output: html_document
date: "2023-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We replicate Table 4a, 5a, and 6a using Doubly Robust (AIPW) estimators. We binarize the treatment $preschool69 \times SSCov$ into an indicator variable which equals 1 if the individual is of preschool age in 1969 and has Sesame Street coverage, but also has above average Sesame Street coverage.

# Load and Prepare Data

```{r}
#Preparing census80
census80 = read.table("C:\\Users\\ttrap\\Downloads\\census80.Rda")

census80 = census80 %>% 
      filter(!is.na(gradeage), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))

#Creating treatment indicator. Equals 1 if preschlcov is above average
census80 = census80 %>% 
  mutate(above_avg_preschlcov = ifelse(preschlcov > mean(preschlcov, na.rm = TRUE), 1, 0))
```

```{r}
#Preparing census70
census70 = read.table("C:\\Users\\ttrap\\Downloads\\census70.Rda")
census70 = census70 %>% 
      filter(!is.na(gradeage), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))

#Creating treatment indicator. Equals 1 if preschlcov is above average
census70 = census70 %>% 
  mutate(above_avg_preschlcov = ifelse(preschlcov > mean(preschlcov, na.rm = TRUE), 1, 0))
```

```{r}
#Preparing census90 and census2000
census90 = read.table("C:\\Users\\ttrap\\Downloads\\census90.Rda")
census90 = census90 %>% 
      filter(!is.na(hsdrop), !is.na(hsgrad), !is.na(anycol), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))

#Creating treatment indicator. Equals 1 if preschlcov is above average
census90 = census90 %>% 
  mutate(above_avg_preschlcov = ifelse(preschlcov > mean(preschlcov, na.rm = TRUE), 1, 0))

census2000 = read.table("C:\\Users\\ttrap\\Downloads\\census2000.Rda")
census2000 = census2000 %>% 
      filter(!is.na(lnhrwage), !is.na(working), !is.na(inpov), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))

#Creating treatment indicator. Equals 1 if preschlcov is above average
census2000 = census2000 %>% 
  mutate(above_avg_preschlcov = ifelse(preschlcov > mean(preschlcov, na.rm = TRUE), 1, 0))
```

# Load Libraries/Import Functions

```{r}
library(tidyverse)
library(plm)
library(haven)
library(estimatr)
library(parallel)
library(pbapply)
library(progress)
library(shiny)
```

```{r}
#From Chapter 12 of the textbook

## Chapter 12.3.1
#z - vector of treatment assignment
#x - matrix of covariates
#y - vector of outcomes of interest
doubly_robust = function(z, y, x, out.family = gaussian, 
                  truncps = c(0, 1))
{
     ## fitted propensity score
     pscore   = glm(z ~ x, family = binomial)$fitted.values
     pscore   = pmax(truncps[1], pmin(truncps[2], pscore))
     
     ## fitted potential outcomes
     outcome1 = glm(y ~ x, weights = z, 
                    family = out.family)$fitted.values
     outcome0 = glm(y ~ x, weights = (1 - z), 
                    family = out.family)$fitted.values
     
     ace.reg  = mean(outcome1 - outcome0) 
     
     ## doubly robust estimator
     res1      = y - outcome1
     res0      = y - outcome0
     r.treat   = mean(z*res1/pscore)
     r.control = mean((1 - z)*res0/(1 - pscore))
     ace.dr    = ace.reg + r.treat - r.control

  return(ace.dr)   
}
```

```{r}
#Function to get a clustered subsample
cluster_subsample <- function(data, cluster_variable, n) {
  
  #Empty dataframe to put samples
  subsample <- data.frame()
  
  #Calculate proportion of each cluster
  cluster_proportions = table(data[[cluster_variable]])/nrow(data)
  
  for (cluster in unique(data[[cluster_variable]])) {
  
  #Calculate cluster size to sample, round to nearest integer
  p_k = cluster_proportions[[as.character(cluster)]]
  cluster_size = round(p_k * n)
  
  #Filter for observations in cluster
  cluster_data <- filter(data, !!sym(cluster_variable) == cluster)
  
  # Sample from the current cluster
  sampled_data <- cluster_data[sample(nrow(cluster_data), cluster_size, replace = TRUE), ]
  
  # Append the sampled data to the subsample
  subsample <- rbind(subsample, sampled_data)
  }
  
  return(subsample)
}
```

We slightly modify the bootstrapping function. 

```{r}
bootstrap_table4_DR = function(census, n) {
    
    #Initialize empty vector to put DR_est
    #1: all, 2: male, 3: female, 4: white, 5: black, 6: hispanic
    coef_vec = numeric(length = 6)
    
    #Take subsample of census
    census_sample = cluster_subsample(census, "dmaindex", n)
    
    #Define groups for regressions
    groups <- list(all = quote(TRUE), male = quote(female == 0), female = quote(female == 1), 
               white = quote(whitenh == 1), black = quote(blacknh == 1), hispanic = quote(hispanic == 1))

    group_names = names(groups)
    
    for (i in seq_along(group_names)) {
      group_name = group_names[i]
      group_condition <- groups[[group_name]]
      
      group_data = census_sample %>%
        filter(!!group_condition)
      
      #Randomly permute above_avg_preschlcov, same as randomly permuting covratehat and recalculating
      group_data[["above_avg_preschlcov"]] = sample(group_data[["above_avg_preschlcov"]])
      
      #Doubly Robust estimation
      #We truncate to 95%
      DR_est = doubly_robust(z = group_data$above_avg_preschlcov, y = group_data$gradeage, 
                             x = as.matrix(group_data[, controls]), truncps = c(0.05, 0.95))
      coef_vec[i] = DR_est
      
    }
  return(coef_vec)
}
```

# Table 4a

## Initial DR Estimates

```{r}
#Defining control variables and empty matrices to store coefficients
controls <- c("momhsdrop2", "momhsgrad2", "livwmom", "blacknh", "othernh", "hispanic", 
              "female", "hspercap", "fsage6", "fsage0_6", "start66", "start67", 
              "start68", "start69", "start70", "start71", "start72", "start73", "start74")
#census4a_DR = matrix(NA, nrow = 4, ncol = 6)
```


```{r}
groups <- list(all = quote(TRUE), male = quote(female == 0), female = quote(female == 1), 
               white = quote(whitenh == 1), black = quote(blacknh == 1), hispanic = quote(hispanic == 1))

group_names = names(groups)

for (i in 1:6) {
  group_name = group_names[i]
  group_condition <- groups[[group_name]]
  
  group_data = filter(census80, !!group_condition)
  
  # Estimate the FE model
  DR_est = doubly_robust(z = group_data$above_avg_preschlcov, y = group_data$gradeage, x = as.matrix(group_data[, controls]), truncps = c(0.05, 0.95))
  
  census4a_DR[1, i] <- DR_est
  census4a_DR[3, i] <- nrow(group_data)
  census4a_DR[4, i] <- mean(group_data$above_avg_preschlcov)
}
```

## Table 4a Bootstrap

```{r}
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps
)

# Suppress warnings for bootstrap_table4
result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table4_DR(census80, n = 30000)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create a single matrix this time

#Indices of DR estimates corresponds to
#1: All, 2: Boys, 3: Girls, 4: White, 5: Black, 6: Hispanic
table4a_DR_mat <- do.call(rbind, result_list)

elapsed_time
```

```{r}
#Set the second row to the standard error vector
table4a_DR_se = apply(table4a_DR_mat, 2, sd)

census4a_DR[2, ] = table4a_DR_se
```


# Table 5a

## Initial DR Estimates

```{r}
#Defining control variables and empty matrices to store coefficients
controls <- c("momhsdrop2", "momhsgrad2", "livwmom", "blacknh", "othernh", "hispanic", 
              "female", "hspercap", "fsage6", "fsage0_6", "start66", "start67", 
              "start68", "start69", "start70", "start71", "start72", "start73", "start74")

#census5a_DR <- matrix(NA, nrow = 4, ncol = 6)
```

```{r}
groups <- list(all = quote(TRUE), male = quote(female == 0), female = quote(female == 1), 
               white = quote(whitenh == 1), black = quote(blacknh == 1), hispanic = quote(hispanic == 1))

group_names = names(groups)

for (i in 1:6) {
  group_name = group_names[i]
  group_condition <- groups[[group_name]]
  
  group_data = filter(census70, !!group_condition)
  
  # Estimate the FE model
  DR_est = doubly_robust(z = group_data$above_avg_preschlcov, y = group_data$gradeage, x = as.matrix(group_data[, controls]), truncps = c(0.05, 0.95))
  
  census5a_DR[1, i] <- DR_est
  census5a_DR[3, i] <- nrow(group_data)
  census5a_DR[4, i] <- mean(group_data$above_avg_preschlcov)
}
```

## Table 5a Bootstrap

```{r}
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps
)

# Suppress warnings for bootstrap_table4
result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table4_DR(census70, n = 30000)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create a single matrix this time

#Indices of DR estimates corresponds to
#1: All, 2: Boys, 3: Girls, 4: White, 5: Black, 6: Hispanic
table5a_DR_mat <- do.call(rbind, result_list)

elapsed_time
```

```{r}
table5a_DR_se = apply(table5a_DR_mat, 2, sd)

census5a_DR[2, ] = table5a_DR_se
```

# Table 6a

## Initial DR Estimates

```{r}
#census6a_DR <- matrix(NA, nrow = 4, ncol = 6)
```

```{r}
controls902000 = c("blacknh", 'othernh', 'hispanic', 
                  'female', 'hspercap', 'start66', 'start67', 'start68', 'start69', 
                  'start70', 'start71', 'start72', 'start73', 'start74')
```

```{r}
outcomes90 = c("hsdrop", "hsgrad", "anycol")

for (i in 1:3) {
  
  #DR Estimate
  DR_est = doubly_robust(z = census90$above_avg_preschlcov, y = census90[[outcomes90[i]]], x = as.matrix(census90[, controls902000]), truncps = c(0.05, 0.95))
  
  census6a_DR[1, i] <- DR_est
  census6a_DR[3, i] <- nrow(census90)
  census6a_DR[4, i] <- mean(census90$above_avg_preschlcov)
}

outcomes2000 = c('lnhrwage', 'working', 'inpov')

for (i in 1:3) {
  
  #DR Estimate
  DR_est = doubly_robust(z = census2000$above_avg_preschlcov, y = census2000[[outcomes2000[i]]], x = as.matrix(census2000[, controls902000]), truncps = c(0.05, 0.95))
  
  census6a_DR[1, i + 3] <- DR_est
  census6a_DR[3, i + 3] <- nrow(census2000)
  census6a_DR[4, i + 3] <- mean(census2000$above_avg_preschlcov)
}
```

## Modified Bootstrap Function

```{r}
bootstrap_table6_DR = function(census, n, outcomes) {
    
    #Initialize empty vector to put DR_est
    #1: outcomes[1], #2: outcomes[2], #3: outcomes[3]
    coef_vec = numeric(length = 3)
    
    #Take subsample of census
    census_sample = cluster_subsample(census, "dmaindex", n)
    
    #Define groups for regressions
    outcomes_iter = outcomes
    
    for (i in 1:3) {
      
      #Randomly permute above_avg_preschlcov, same as randomly permuting covratehat and recalculating
      census_sample[["above_avg_preschlcov"]] = sample(census_sample[["above_avg_preschlcov"]])
      
      #Doubly Robust estimation
      #We truncate to 95%
      DR_est = doubly_robust(z = census_sample$above_avg_preschlcov, y = census_sample[[outcomes_iter[i]]], 
                             x = as.matrix(census_sample[, controls902000]), truncps = c(0.05, 0.95))
      coef_vec[i] = DR_est
      
    }
  return(coef_vec)
}
```

## census90 Bootstrap

```{r}
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps
)

# Suppress warnings for bootstrap_table4
result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table6_DR(census90, n = 30000, outcomes90)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create a single matrix this time

#Indices of DR estimates corresponds to
#1: All, 2: Boys, 3: Girls, 4: White, 5: Black, 6: Hispanic
table6a_DR_mat_90 <- do.call(rbind, result_list)

elapsed_time
```

```{r}
table6a_DR_se_90 = apply(table6a_DR_mat_90, 2, sd)

census6a_DR[2, 1:3] = table6a_DR_se_90
```

## census2000 Bootstrap

```{r}
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps
)

# Suppress warnings for bootstrap_table4
result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table6_DR(census2000, n = 30000, outcomes2000)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create a single matrix this time

#Indices of DR estimates corresponds to
#1: All, 2: Boys, 3: Girls, 4: White, 5: Black, 6: Hispanic
table6a_DR_mat_2000 <- do.call(rbind, result_list)

elapsed_time
```

```{r}
table6a_DR_se_2000 = apply(table6a_DR_mat_2000, 2, sd)

census6a_DR[2, 4:6] = table6a_DR_se_2000
```

# Making LaTeX Tables

## Table 4a

```{r}
census4a_table_DR <- xtable(census4a_DR, digits = 3)

# Print LaTeX code
print(census4a_table_DR, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

## Table 5a

```{r}
census5a_table_DR <- xtable(census5a_DR, digits = 3)

# Print LaTeX code
print(census5a_table_DR, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

## Table 6a

```{r}
census6a_table_DR <- xtable(census6a_DR, digits = 3)

# Print LaTeX code
print(census6a_table_DR, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```





































































