---
title: "STAT 156 Table Replication (Alternative Bootstrap)"
output: html_document
date: "2023-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(tidyverse)
library(plm)
library(haven)
library(estimatr)
library(parallel)
library(pbapply)
library(progress)
library(shiny)
```

# Bootstrap compromise

Due to computational restraints, we have to do subsample clustered bootstrapping. Although there may be asymptotic results published somewhere, the results are unknown to us, and are somewhat out of scope of this replication. Nonetheless, we proceed with the following algorithm:

1.  Choose a total sample size $n$ (we will choose $n = 10,000$).
2.  Calculate the proportions of data points in each cluster, denoted as $p_k$.
3.  For each cluster $k$, randomly sample with replacement $p_kn$ data points, rounding up to the nearest integer.
4.  Proceed with the bootstrapping/estimation procedure as usual.


Since we cannot exactly replicate the bootstrap procedure due to computational restraints, we will attempt to replicate some degree of randomness coming from the coverage simulation by randomly permuting the simulated coverage within bootstrap sample, then calculating coverage variables.

```{r}
#Function to get a clustered subsample
cluster_subsample <- function(data, cluster_variable, n) {
  
  #Empty dataframe to put samples
  subsample <- data.frame()
  
  #Calculate proportion of each cluster
  cluster_proportions = table(data[[cluster_variable]])/nrow(data)
  
  for (cluster in unique(data[[cluster_variable]])) {
  
  #Calculate cluster size to sample, round to nearest integer
  p_k = cluster_proportions[[as.character(cluster)]]
  cluster_size = round(p_k * n)
  
  #Filter for observations in cluster
  cluster_data <- filter(data, !!sym(cluster_variable) == cluster)
  
  # Sample from the current cluster
  sampled_data <- cluster_data[sample(nrow(cluster_data), cluster_size, replace = TRUE), ]
  
  # Append the sampled data to the subsample
  subsample <- rbind(subsample, sampled_data)
  }
  
  return(subsample)
}
```


# Table 4

```{r}
#Preparing census80
census80 = read.table("C:\\Users\\ttrap\\Downloads\\census80.Rda")

census80 = census80 %>% 
      filter(!is.na(gradeage), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))
```

### Initial Regression Results

```{r}
#Defining control variables and empty matrices to store coefficients
controls <- c("momhsdrop2", "momhsgrad2", "livwmom", "blacknh", "othernh", "hispanic", 
              "female", "hspercap", "fsage6", "fsage0_6", "start66", "start67", 
              "start68", "start69", "start70", "start71", "start72", "start73", "start74")

census4a <- matrix(NA, nrow = 4, ncol = 6)
census4b <- matrix(NA, nrow = 9, ncol = 6)
```

```{r}
groups <- list(all = quote(TRUE), male = quote(female == 0), female = quote(female == 1), 
               white = quote(whitenh == 1), black = quote(blacknh == 1), hispanic = quote(hispanic == 1))

group_names = names(groups)

for (i in 1:6) {
  group_name = group_names[i]
  group_condition <- groups[[group_name]]
  
  group_data = filter(census80, !!group_condition)
  
  # Estimate the FE model
  model_1 <- plm(as.formula(paste("gradeage ~ preschlcov +", paste(controls, collapse = " + "))), 
                   data = group_data, index = c("statecty", "statecohort"), model = "within")
  
  census4a[1, i] <- coef(model_1)["preschlcov"]
  census4a[3, i] <- nrow(group_data)
  census4a[4, i] <- mean(group_data$gradeage)
  
  model_2 <- plm(as.formula(paste("gradeage ~ covst6768 + covst69 + covst7072 + covst7374 + covratehat +", paste(controls, collapse = " + "))), 
                 data = group_data, index = c("statecty", "statecohort"), model = "within")
  
  census4b[1, i] <- coef(model_2)["covst6768"]
  census4b[3, i] <- coef(model_2)["covst69"]
  census4b[5, i] <- coef(model_2)["covst7072"]
  census4b[7, i] <- coef(model_2)["covst7374"]
  census4b[9, i] <- nrow(group_data)
}
```

### Table 4 Bootstrap

```{r}
bootstrap_table4 = function(census, n) {
    
    #Initialize empty list to put coef_vecs
    #1: all, 2: male, 3: female, 4: white, 5: black, 6: hispanic
    coef_list = list()
    
    #Take subsample of census
    census_sample = cluster_subsample(census, "dmaindex", n)
    
    #Define groups for regressions
    groups <- list(all = quote(TRUE), male = quote(female == 0), female = quote(female == 1), 
               white = quote(whitenh == 1), black = quote(blacknh == 1), hispanic = quote(hispanic == 1))

    group_names = names(groups)
    
    for (i in 1:6) {
      group_name = group_names[i]
      group_condition <- groups[[group_name]]
      
      group_data = census_sample %>%
        filter(!!group_condition)
      
      #Randomly permute covratehat
      group_data[["covratehat"]] = sample(group_data[["covratehat"]])
    
      #Recalculate coverage variables
      years <- 65:74
      for (yr in years) {
      start_col <- paste("start", yr, sep = "")
      covstyr_col <- paste("covst", yr, sep = "")
      group_data[[covstyr_col]] <- group_data$covratehat * group_data[[start_col]]
      }
    
      group_data = group_data %>% 
        mutate(covst6768 = covratehat*start6768, 
               covst7072 = covratehat*start7072,
               covst7374 = covratehat*start7374)
      
      
      #Estimate regressions
      model_1 <- plm(as.formula(paste("gradeage ~ preschlcov +", paste(controls, collapse = " + "))), 
                   data = group_data, index = c("statecty", "statecohort"), model = "within")
      
      model_2 <- plm(as.formula(paste("gradeage ~ covst6768 + covst69 + covst7072 + covst7374 + covratehat +", paste(controls, collapse = " + "))), data = group_data, index = c("statecty", "statecohort"), model = "within")
  
      coef_vec = c(coef(model_1)["preschlcov"][1], coef(model_2)["covst6768"][1], coef(model_2)["covst69"][1], 
               coef(model_2)["covst7072"][1], coef(model_2)["covst7374"][1])
      
      #Add coef_vec to list
      coef_list[[i]] = coef_vec
      
    }
  return(coef_list)
}
```

```{r}
#Test run, it works!
#bootstrap_table4(census80, 20000)
```

```{r}
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps
)

# Suppress warnings for bootstrap_table4
result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table4(census80, n = 30000)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create matrices

#1: All, 2: Boys, 3: Girls, 4: White, 5: Black, 6: Hispanic
combined_matrices_table4 <- lapply(seq_along(result_list[[1]]), function(j) {
     do.call(rbind, lapply(result_list, function(rep_result) rep_result[[j]]))
})

elapsed_time
```

```{r}
#Create vectors of bootstrapped S.E. to put into census4a and census4b

std_dev_list_table4 <- lapply(combined_matrices_table4, function(mat) apply(mat, 2, sd))

std_dev_matrix_table4 <- do.call(rbind, std_dev_list_table4)

std_dev_matrix_table4
```

```{r}
#Setting second row of census4a to the 1st column of std_dev_matrix_table4
census4a[2, ] = std_dev_matrix_table4[, 1]

#Setting all empty rows in census4b to respective std_dev columns
census4b[2, ] = std_dev_matrix_table4[, 2]
census4b[4, ] = std_dev_matrix_table4[, 3]
census4b[6, ] = std_dev_matrix_table4[, 4]
census4b[8, ] = std_dev_matrix_table4[, 5]
```


# Table 5

The process for table 5 follows the same process as table 4. The only thing that changes is the data. 

```{r}
census70 = read.table("C:\\Users\\ttrap\\Downloads\\census70.Rda")
census70 = census70 %>% 
      filter(!is.na(gradeage), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))
```

### Initial Regression Results

```{r}
#Defining control variables and empty matrices to store coefficients
controls <- c("momhsdrop2", "momhsgrad2", "livwmom", "blacknh", "othernh", "hispanic", 
              "female", "hspercap", "fsage6", "fsage0_6", "start66", "start67", 
              "start68", "start69", "start70", "start71", "start72", "start73", "start74")

census5a <- matrix(NA, nrow = 4, ncol = 6)
census5b <- matrix(NA, nrow = 9, ncol = 6)
```

```{r}
groups <- list(all = quote(TRUE), male = quote(female == 0), female = quote(female == 1), 
               white = quote(whitenh == 1), black = quote(blacknh == 1), hispanic = quote(hispanic == 1))

group_names = names(groups)

for (i in 1:6) {
  group_name = group_names[i]
  group_condition <- groups[[group_name]]
  
  group_data = filter(census70, !!group_condition)
  
  # Estimate the FE model
  model_1 <- plm(as.formula(paste("gradeage ~ preschlcov +", paste(controls, collapse = " + "))), 
                   data = group_data, index = c("statecty", "statecohort"), model = "within")
  
  census5a[1, i] <- coef(model_1)["preschlcov"]
  census5a[3, i] <- nrow(group_data)
  census5a[4, i] <- mean(group_data$gradeage)
  
  model_2 <- plm(as.formula(paste("gradeage ~ covst6768 + covst69 + covst7072 + covst7374 + covratehat +", paste(controls, collapse = " + "))), 
                 data = group_data, index = c("statecty", "statecohort"), model = "within")
  
  census5b[1, i] <- coef(model_2)["covst6768"]
  census5b[3, i] <- coef(model_2)["covst69"]
  census5b[5, i] <- coef(model_2)["covst7072"]
  census5b[7, i] <- coef(model_2)["covst7374"]
  census5b[9, i] <- nrow(group_data)
}
```

### Table 5 Bootstrap

```{r}
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps
)

# Suppress warnings for bootstrap_table4
result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table4(census70, n = 30000)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create matrices

#1: All, 2: Boys, 3: Girls, 4: White, 5: Black, 6: Hispanic
combined_matrices_table5 <- lapply(seq_along(result_list[[1]]), function(j) {
     do.call(rbind, lapply(result_list, function(rep_result) rep_result[[j]]))
})

elapsed_time
```

```{r}
#Create vectors of bootstrapped S.E. to put into census5a and census5b

std_dev_list_table5 <- lapply(combined_matrices_table5, function(mat) apply(mat, 2, sd))

std_dev_matrix_table5 <- do.call(rbind, std_dev_list_table5)

std_dev_matrix_table5
```

```{r}
#Setting second row of census4a to the 1st column of std_dev_matrix_table4
census5a[2, ] = std_dev_matrix_table5[, 1]

#Setting all empty rows in census4b to respective std_dev columns
census5b[2, ] = std_dev_matrix_table5[, 2]
census5b[4, ] = std_dev_matrix_table5[, 3]
census5b[6, ] = std_dev_matrix_table5[, 4]
census5b[8, ] = std_dev_matrix_table5[, 5]
```


# Table 6

The process for Table 6 is slightly different, so we will slightly modify the process, since Table 6 uses 2 datasets, and runs two separate regression processes.

Table 6 does not filter by groups: instead, it estimates effects on multiple outcomes.
Table 6 also uses slightly different controls. 

```{r}
#First 3 columns are results for census90, last 3 are for census2000
census6a <- matrix(NA, nrow = 4, ncol = 6)
census6b <- matrix(NA, nrow = 9, ncol = 6)
```

## Loading Data

```{r}
census90 = read.table("C:\\Users\\ttrap\\Downloads\\census90.Rda")
census90 = census90 %>% 
      filter(!is.na(hsdrop), !is.na(hsgrad), !is.na(anycol), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))

census2000 = read.table("C:\\Users\\ttrap\\Downloads\\census2000.Rda")
census2000 = census2000 %>% 
      filter(!is.na(lnhrwage), !is.na(working), !is.na(inpov), !is.na(preschlcov), !is.na(presch169),
           !is.na(covratehat), !is.na(covst6768), !is.na(covst69), !is.na(covst7072), !is.na(covst7374))
```

### Initial Regression Results

```{r}
controls902000 = c("blacknh", 'othernh', 'hispanic',
                  'female', 'hspercap', 'start66', 'start67', 'start68', 'start69', 
                  'start70', 'start71', 'start72', 'start73', 'start74')
```

i from 1-3 uses census90, i from 4-6 uses census2000

```{r}
outcomes90 = c("hsdrop", "hsgrad", "anycol")

for (i in 1:3) {
  
  # Estimate the FE model
  model_1 <- plm(as.formula(paste(outcomes90[i], paste("~ preschlcov +", paste(controls902000, collapse = " + ")))), data = census90, index = c("statecty", "statecohort"), model = "within")
  
  census6a[1, i] <- coef(model_1)["preschlcov"]
  census6a[3, i] <- nrow(census90)
  census6a[4, i] <- mean(census90[[outcomes90[i]]])
  
  model_2 <- plm(as.formula(paste(outcomes90[i], paste("~ covst6768 + covst69 + covst7072 + covst7374 + covratehat +", paste(controls902000, collapse = " + ")))), data = census90, index = c("statecty", "statecohort"), model = "within")
  
  census6b[1, i] <- coef(model_2)["covst6768"]
  census6b[3, i] <- coef(model_2)["covst69"]
  census6b[5, i] <- coef(model_2)["covst7072"]
  census6b[7, i] <- coef(model_2)["covst7374"]
  census6b[9, i] <- nrow(census90)
}

outcomes2000 = c('lnhrwage', 'working', 'inpov')

for (i in 1:3) {
  
  # Estimate the FE model
  model_1 <- plm(as.formula(paste(outcomes2000[i], paste("~ preschlcov +", paste(controls902000, collapse = " + ")))), data = census2000, index = c("statecty", "statecohort"), model = "within")
  
  census6a[1, i + 3] <- coef(model_1)["preschlcov"]
  census6a[3, i + 3] <- nrow(census2000)
  census6a[4, i + 3] <- mean(census2000[[outcomes2000[i]]])
  
  model_2 <- plm(as.formula(paste(outcomes2000[i], paste("~ covst6768 + covst69 + covst7072 + covst7374 + covratehat +", paste(controls902000, collapse = " + ")))), data = census2000, index = c("statecty", "statecohort"), model = "within")
  
  census6b[1, i + 3] <- coef(model_2)["covst6768"]
  census6b[3, i + 3] <- coef(model_2)["covst69"]
  census6b[5, i + 3] <- coef(model_2)["covst7072"]
  census6b[7, i + 3] <- coef(model_2)["covst7374"]
  census6b[9, i + 3] <- nrow(census2000)
}
```

### Table 6 Bootstrap

This bootstrap is slightly different, since there are slightly different variables. Most importantly, since there are not separate groups to consider, we randomly permute covratehat across the whole census data before we take a sample. 

```{r}
#Returns a list of 3 vectors, each with 5 coefficients
#Need to change this so it returns a list of 6 vectors, each with 5 coefficients
#Meaning that we do both census bootstraps within this function

bootstrap_table6 = function(census, n, outcomes) {
    
    #Initialize empty list to put coef_vecs
    #1: all, 2: male, 3: female, 4: white, 5: black, 6: hispanic
    coef_list = list()
    
    #Randomly permute covratehat before taking sample
    census[["covratehat"]] = sample(census[["covratehat"]])
    
    #Take subsample of census
    census_sample = cluster_subsample(census, "dmaindex", n)
    
    #Define groups for regressions
    outcomes_iter = outcomes
    
    for (i in seq_along(outcomes_iter)) {
    
      #Recalculate coverage variables
      years <- 65:74
      for (yr in years) {
      start_col <- paste("start", yr, sep = "")
      covstyr_col <- paste("covst", yr, sep = "")
      census_sample[[covstyr_col]] <- census_sample$covratehat * census_sample[[start_col]]
      }
    
      census_sample = census_sample %>% 
        mutate(covst6768 = covratehat*start6768, 
               covst7072 = covratehat*start7072,
               covst7374 = covratehat*start7374)
      
      
      #Estimate regressions
      model_1 <- plm(as.formula(paste(outcomes_iter[i], paste("~ preschlcov +", paste(controls902000, collapse = " + ")))), data = census_sample, index = c("statecty", "statecohort"), model = "within")
      
      model_2 <- plm(as.formula(paste(outcomes_iter[i], paste("~ covst6768 + covst69 + covst7072 + covst7374 + covratehat +", paste(controls902000, collapse = " + ")))), data = census_sample, index = c("statecty", "statecohort"), model = "within")
  
      coef_vec = c(coef(model_1)["preschlcov"][1], coef(model_2)["covst6768"][1], coef(model_2)["covst69"][1], 
               coef(model_2)["covst7072"][1], coef(model_2)["covst7374"][1])
      
      #Add coef_vec to list
      coef_list[[i]] = coef_vec
      
    }
  return(coef_list)
}
```

Note that now, we work with two lists of 3 vectors each: one list will correspond to census90, the other corresponds to census2000.

### census90 Bootstrap

```{r}
# Set a seed for reproducibility
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps * 2  # Update total for two datasets
)


# Suppress warnings for pbreplicate
# Suppress warnings for bootstrap_table4

result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table6(census90, n = 30000, outcomes90)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create matrices

#Order of coefficient matrices:
#1: hsdrop, 2: hsgrad, 3: anycol
combined_matrices_table6_90 <- lapply(seq_along(result_list[[1]][1:3]), function(j) {
     do.call(rbind, lapply(result_list, function(rep_result) rep_result[[j]]))
})

elapsed_time
```

```{r}
std_dev_list_table6_90 <- lapply(combined_matrices_table6_90, function(mat) apply(mat, 2, sd))

std_dev_matrix_table6_90 <- do.call(rbind, std_dev_list_table6_90)

std_dev_matrix_table6_90
```

```{r}
census6a[2, 1:3] = std_dev_matrix_table6_90[, 1]

census6b[2, 1:3] = std_dev_matrix_table6_90[, 2]
census6b[4, 1:3] = std_dev_matrix_table6_90[, 3]
census6b[6, 1:3] = std_dev_matrix_table6_90[, 4]
census6b[8, 1:3] = std_dev_matrix_table6_90[, 5]
```


### census2000 Bootstrap

```{r}
# Set a seed for reproducibility
set.seed(123)

# Set the number of repetitions
reps <- 200

# Timing the execution of the code
start_time <- Sys.time()

# Create a progress bar
pb <- progress_bar$new(
  format = "[:bar] :percent Elapsed: :elapsed Time Remaining: :eta",
  total = reps * 2  # Update total for two datasets
)


# Suppress warnings for pbreplicate
# Suppress warnings for bootstrap_table4

result_list <- pbreplicate(
  reps,
  suppressWarnings(bootstrap_table6(census2000, n = 30000, outcomes2000)),
  simplify = FALSE
)

# Timing the end of the code
end_time <- Sys.time()

# Calculate the elapsed time
elapsed_time <- end_time - start_time

# Combine the vectors across coefficients to create matrices

#Order of coefficient matrices:
#1: hsdrop, 2: hsgrad, 3: anycol, 4: lnhrwage, 5: working, 6: inpov
combined_matrices_table6_2000 <- lapply(seq_along(result_list[[1]]), function(j) {
     do.call(rbind, lapply(result_list, function(rep_result) rep_result[[j]]))
})

elapsed_time
```

```{r}
#Create vectors of bootstrapped S.E. to put into census5a and census5b

std_dev_list_table6_2000 <- lapply(combined_matrices_table6_2000, function(mat) apply(mat, 2, sd))

std_dev_matrix_table6_2000 <- do.call(rbind, std_dev_list_table6_2000)

std_dev_matrix_table6_2000
```

```{r}
census6a[2, 4:6] = std_dev_matrix_table6_2000[, 1]

census6b[2, 4:6] = std_dev_matrix_table6_2000[, 2]
census6b[4, 4:6] = std_dev_matrix_table6_2000[, 3]
census6b[6, 4:6] = std_dev_matrix_table6_2000[, 4]
census6b[8, 4:6] = std_dev_matrix_table6_2000[, 5]
```


# Converting Matrices into LaTeX Tables

```{r}
library(xtable)
library(stargazer)
```

### Table 4

```{r}
# Convert matrix to LaTeX table
census4a_table <- xtable(census4a, digits = 3)

# Print LaTeX code
print(census4a_table, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

```{r}
# Convert matrix to LaTeX table
census4b_table <- xtable(census4b, digits = 3)

# Print LaTeX code
print(census4b_table, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

```{r}
stargazer(census4b, title = "My Matrix Table", label = "tab:my_matrix", summary = FALSE)
```

### Table 5

```{r}
# Convert matrix to LaTeX table
census5a_table <- xtable(census5a, digits = 3)

# Print LaTeX code
print(census5a_table, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

```{r}
# Convert matrix to LaTeX table
census5b_table <- xtable(census5b, digits = 3)

# Print LaTeX code
print(census5b_table, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

### Table 6

```{r}
census6a_table <- xtable(census6a, digits = 3)

# Print LaTeX code
print(census6a_table, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```

```{r}
census6b_table <- xtable(census6b, digits = 3)

# Print LaTeX code
print(census6b_table, include.rownames = TRUE, caption.placement = "top", booktabs = TRUE)
```



































